\textbf{skater}: An R package for SNP-based Kinship Analysis, Testing, and Evaluation

### 2021-11-29 paper/paper.tex
updated first draft of f1000research paper

@@

The skater package has several functions for importing, parsing, and analyzing pedigree data. Pedigrees define familial relationships in a hierarchical structure. Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit. The skater package contains the function \texttt{read\_fam()} to read in a PLINK-formatted .fam file and another function \texttt{fam2ped()} to convert the content into a pedigree object as a nested tibble with one row per family. All pedigree processing from skater internally leverages a data structure from the kinship2 package \citep{sinnwell2014}. Further functions such as \texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of the pedigree for each family, while \texttt{ped2kinpair()} produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).

@@

The \texttt{dibble()} function creates a \textbf{d}egree \textbf{i}nference t\textbf{ibble}, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in \citet{manichaikul2010}. The \texttt{kin2degree()} function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness. When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method. The skater package adapts a \texttt{confusion\_matrix()} function from \citet{clark2021} to provide standard contingency table metrics (e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more thoroughly in the skater package vignette and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation. In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number. Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).

@@

Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020} detect shared IBD segments between individuals. The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and converts that shared cM to a kinship coefficient. In addition to inferred segments, these functions can estimate ``truth'' kinship from simulated IBD segments \citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise IBD segments from IBD inference tools and from simulated IBD segments. The \texttt{read\_map()} function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the \texttt{ibd2kin()} function.

>>>>

The \texttt{skater} package provides a collection of analysis and utility functions for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation as an \textbf{R} package. Functions in the package include tools for working with pedigree data, performing relationship degree inference, assessing classification accuracy, and summarizing IBD segment data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(skater)}
\end{Highlighting}
\end{Shaded}

@@

Pedigrees define familial relationships in a hierarchical structure. One of the common formats used by PLINK \citep{purcell2007} and other genetic analysis tools is the \texttt{.fam} file. A \texttt{.fam} file is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit. The package includes \texttt{read\_fam()} to read files in this format:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{famfile }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\AttributeTok{package=}\StringTok{"skater"}\NormalTok{, }\AttributeTok{mustWork=}\ConstantTok{TRUE}\NormalTok{)}
\NormalTok{fam }\OtherTok{\textless{}{-}} \FunctionTok{read\_fam}\NormalTok{(famfile)}
\NormalTok{fam}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 64 x 6
##    fid      id                dadid             momid               sex affected
##    <chr>    <chr>             <chr>             <chr>             <int>    <int>
##  1 testped1 testped1_g1-b1-s1 0                 0                     1        1
##  2 testped1 testped1_g1-b1-i1 0                 0                     2        1
##  3 testped1 testped1_g2-b1-s1 0                 0                     1        1
##  4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_g1-b1-i1     2        1
##  5 testped1 testped1_g2-b2-s1 0                 0                     1        1
##  6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_g1-b1-i1     2        1
##  7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_g2-b1-i1     2        1
##  8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_g2-b2-i1     1        1
##  9 testped2 testped2_g1-b1-s1 0                 0                     2        1
## 10 testped2 testped2_g1-b1-i1 0                 0                     1        1
## # ... with 54 more rows
\end{verbatim}

Family structures imported from \texttt{.fam} formated files can then be translated to the \texttt{pedigree} structure used by the \texttt{kinship2} package \citep{sinnwell2014}. The ``fam'' format may include multiple families, and the \texttt{fam2ped()} function will collapse them all into a \texttt{tibble} with one row per family:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds }\OtherTok{\textless{}{-}} \FunctionTok{fam2ped}\NormalTok{(fam)}
\NormalTok{peds}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 8 x 3
##   fid      data             ped       
##   <chr>    <list>           <list>    
## 1 testped1 <tibble [8 x 5]> <pedigree>
## 2 testped2 <tibble [8 x 5]> <pedigree>
## 3 testped3 <tibble [8 x 5]> <pedigree>
## 4 testped4 <tibble [8 x 5]> <pedigree>
## 5 testped5 <tibble [8 x 5]> <pedigree>
## 6 testped6 <tibble [8 x 5]> <pedigree>
## 7 testped7 <tibble [8 x 5]> <pedigree>
## 8 testped8 <tibble [8 x 5]> <pedigree>
\end{verbatim}

In the example above, the resulting \texttt{tibble} is nested by family ID. The \texttt{data} column contains the individual family information, while the \texttt{ped} column contains the pedigree object for that family. Using standard tidyverse operations, the resulting tibble can be unnested for any particular family:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(fid}\SpecialCharTok{==}\StringTok{"testped1"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 8 x 7
##   fid      id                dadid             momid        sex affected ped    
##   <chr>    <chr>             <chr>             <chr>      <int>    <dbl> <list> 
## 1 testped1 testped1_g1-b1-s1 <NA>              <NA>           1        1 <pedig~
## 2 testped1 testped1_g1-b1-i1 <NA>              <NA>           2        1 <pedig~
## 3 testped1 testped1_g2-b1-s1 <NA>              <NA>           1        1 <pedig~
## 4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 5 testped1 testped1_g2-b2-s1 <NA>              <NA>           1        1 <pedig~
## 6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_~     2        1 <pedig~
## 8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_~     1        1 <pedig~
\end{verbatim}

A single pedigree can also be inspected or visualized (standard base R plot arguments such as \texttt{mar} or \texttt{cex} can be used to adjust aesthetics):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Pedigree object with 8 subjects
## Bit size= 4
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{mar=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\AttributeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{paper_files/figure-latex/plotped-1} \end{center}

The \texttt{plot\_pedigree()} function from \texttt{skater} will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from family:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_pedigree}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped, }\AttributeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The \texttt{ped2kinpair()} function takes a pedigree object and produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.

The function can be run on a single family:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ped2kinpair}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 36 x 3
##    id1               id2                   k
##    <chr>             <chr>             <dbl>
##  1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5  
##  2 testped1_g1-b1-i1 testped1_g1-b1-s1 0    
##  3 testped1_g1-b1-s1 testped1_g2-b1-s1 0    
##  4 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25 
##  5 testped1_g1-b1-s1 testped1_g2-b2-s1 0    
##  6 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25 
##  7 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125
##  8 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125
##  9 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5  
## 10 testped1_g1-b1-i1 testped1_g2-b1-s1 0    
## # ... with 26 more rows
\end{verbatim}

This function can also be mapped over all families in the pedigree:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{kinpairs }\OtherTok{\textless{}{-}} 
\NormalTok{  peds }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{pairs=}\NormalTok{purrr}\SpecialCharTok{::}\FunctionTok{map}\NormalTok{(ped, ped2kinpair)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(fid, pairs) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{pairs)}
\NormalTok{kinpairs}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 288 x 4
##    fid      id1               id2                   k
##    <chr>    <chr>             <chr>             <dbl>
##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5  
##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0    
##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0    
##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25 
##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0    
##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25 
##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125
##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125
##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5  
## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0    
## # ... with 278 more rows
\end{verbatim}

Note that this maps \texttt{ped2kinpair()} over all \texttt{ped} objects in the input \texttt{tibble}, and that relationships are not shown for between-family relationships.

@@

The \texttt{dibble()} function creates a \textbf{d}egree \textbf{i}nference \texttt{tibble}, with degrees up to the specified \texttt{max\_degree} (default=3), expected kinship coefficient, and lower (\texttt{l}) and upper (\texttt{u}) inference ranges as defined in \citet{manichaikul2010}. Degree 0 corresponds to self / identity / monozygotic twins, with an expected kinship coefficient of 0.5, with inference range \textgreater=0.354. Anything beyond the maximum degree resolution is considered unrelated (degree \texttt{NA}).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dibble}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 5 x 4
##   degree      k       l      u
##    <int>  <dbl>   <dbl>  <dbl>
## 1      0 0.5     0.354  1     
## 2      1 0.25    0.177  0.354 
## 3      2 0.125   0.0884 0.177 
## 4      3 0.0625  0.0442 0.0884
## 5     NA 0      -1      0.0442
\end{verbatim}

The degree inference \texttt{max\_degree} default is 3. Change this argument to allow more granular degree inference ranges:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dibble}\NormalTok{(}\AttributeTok{max\_degree =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 7 x 4
##   degree      k       l      u
##    <int>  <dbl>   <dbl>  <dbl>
## 1      0 0.5     0.354  1     
## 2      1 0.25    0.177  0.354 
## 3      2 0.125   0.0884 0.177 
## 4      3 0.0625  0.0442 0.0884
## 5      4 0.0312  0.0221 0.0442
## 6      5 0.0156  0.0110 0.0221
## 7     NA 0      -1      0.0110
\end{verbatim}

Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant. The \texttt{dibble()} function will emit a warning with \texttt{max\_degree} \textgreater=10, and will stop with an error at \textgreater=12.

The \texttt{kin2degree()} function infers the relationship degree given a kinship coefficient and a \texttt{max\_degree} up to which anything more distant is treated as unrelated. Example first degree relative:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1
\end{verbatim}

Example 4th degree relative, but using the default max\_degree resolution of 3:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] NA
\end{verbatim}

Example 4th degree relative, but increasing the degree resolution:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 4
\end{verbatim}

The \texttt{kin2degree()} function is vectorized over values of \texttt{k}, so it can be used inside of a \texttt{mutate} on a \texttt{tibble} of kinship coefficients:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get two pairs from each type of relationship we have in kinpairs:}
\NormalTok{kinpairs\_subset }\OtherTok{\textless{}{-}} 
\NormalTok{  kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{group\_by}\NormalTok{(k) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{slice}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{)}
\NormalTok{kinpairs\_subset}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 10 x 4
## # Groups:   k [5]
##    fid      id1               id2                    k
##    <chr>    <chr>             <chr>              <dbl>
##  1 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     
##  2 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     
##  3 testped1 testped1_g3-b1-i1 testped1_g3-b2-i1 0.0625
##  4 testped2 testped2_g3-b1-i1 testped2_g3-b2-i1 0.0625
##  5 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 
##  6 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 
##  7 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  
##  8 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  
##  9 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   
## 10 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Infer degree out to third degree relatives:}
\NormalTok{kinpairs\_subset }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 10 x 5
## # Groups:   k [5]
##    fid      id1               id2                    k degree
##    <chr>    <chr>             <chr>              <dbl>  <int>
##  1 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0          NA
##  2 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0          NA
##  3 testped1 testped1_g3-b1-i1 testped1_g3-b2-i1 0.0625      3
##  4 testped2 testped2_g3-b1-i1 testped2_g3-b2-i1 0.0625      3
##  5 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125       2
##  6 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125       2
##  7 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25        1
##  8 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25        1
##  9 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5         0
## 10 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5         0
\end{verbatim}

\hypertarget{benchmarking-degree-classification}{%
\subsection{Benchmarking Degree Classification}\label{benchmarking-degree-classification}}

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth. When aggregated over many relationships and inferences, this method can help benchmark performance of a particular kinship analysis method.

The skater package adapts a \texttt{confusion\_matrix()} function from \citet{clark2021} to provide standard contingency table metrics (e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric. The \texttt{confusion\_matrix()} function on its own outputs a list with four objects:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A \texttt{tibble} with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs.~the guessing rate.
\item
  A \texttt{tibble} with contingency table statistics calculated for each class. Details on the statistics calculated for each class can be reviewed on the help page for \texttt{?confusion\_matrix}.
\item
  A \texttt{matrix} with the contingency table object itself.
\item
  A \texttt{vector} with the reciprocal RMSE (R-RMSE). The R-RMSE is calculated as \texttt{sqrt(mean((1/(Target+.5)-1/(Predicted+.5))\^{}2)))}, and represents an alternative to classification accuracy when benchmarking relationship degree estimation. Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree). The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.
\end{enumerate}

To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Function to randomly flip levels of a factor (at 20\%, by default)}
\NormalTok{randomflip }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, }\AttributeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{runif}\NormalTok{(}\FunctionTok{length}\NormalTok{(x))}\SpecialCharTok{\textless{}}\NormalTok{p, }\FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(x)), x)}

\CommentTok{\# Infer degree (truth/target) using kin2degree, then randomly flip 20\% of them}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
\NormalTok{kinpairs\_inferred }\OtherTok{\textless{}{-}}\NormalTok{ kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\NormalTok{tidyr}\SpecialCharTok{::}\FunctionTok{replace\_na}\NormalTok{(degree\_truth, }\StringTok{"unrelated"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_inferred=}\FunctionTok{randomflip}\NormalTok{(degree\_truth))}
\NormalTok{kinpairs\_inferred}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 288 x 6
##    fid      id1               id2                   k degree_truth degree_inferred
##    <chr>    <chr>             <chr>             <dbl> <chr>        <chr>          
##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   0            0              
##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     unrelated    unrelated      
##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     unrelated    unrelated      
##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  1            1              
##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0     unrelated    unrelated      
##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  1            1              
##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 2            2              
##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 2            1              
##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5   0            0              
## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0     unrelated    unrelated      
## # ... with 278 more rows
\end{verbatim}

Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $Accuracy
## # A tibble: 1 x 5
##   Accuracy `Accuracy LL` `Accuracy UL` `Accuracy Guessing` `Accuracy P-value`
##      <dbl>         <dbl>         <dbl>               <dbl>              <dbl>
## 1    0.812         0.763         0.856               0.333           1.09e-62
## 
## $Other
## # A tibble: 6 x 15
##   Class         N `Sensitivity/Re~ `Specificity/TN~ `PPV/Precision`   NPV `F1/Dice`
##   <chr>     <dbl>            <dbl>            <dbl>           <dbl> <dbl>     <dbl>
## 1 0          64              0.75             0.964           0.857 0.931     0.8  
## 2 1          72              0.806            0.944           0.829 0.936     0.817
## 3 2          48              0.833            0.967           0.833 0.967     0.833
## 4 3           8              0.75             0.936           0.25  0.992     0.375
## 5 unrelated  96              0.854            0.958           0.911 0.929     0.882
## 6 Average    57.6            0.799            0.954           0.736 0.951     0.741
## # ... with 8 more variables: Prevalence <dbl>, Detection Rate <dbl>,
## #   Detection Prevalence <dbl>, Balanced Accuracy <dbl>, FDR <dbl>, FOR <dbl>,
## #   FPR/Fallout <dbl>, FNR <dbl>
## 
## $Table
##            Target
## Predicted    0  1  2  3 unrelated
##   0         48  4  2  1         1
##   1          5 58  4  0         3
##   2          0  3 40  1         4
##   3          8  4  0  6         6
##   unrelated  3  3  2  0        82
## 
## $recip_rmse
## [1] 0.4665971
\end{verbatim}

Standard tidyverse functions such as \texttt{purrr::pluck()} can be used to isolate just the contingency table:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            Target
## Predicted    0  1  2  3 unrelated
##   0         48  4  2  1         1
##   1          5 58  4  0         3
##   2          0  3 40  1         4
##   3          8  4  0  6         6
##   unrelated  3  3  2  0        82
\end{verbatim}

The \texttt{confusion\_matrix()} function includes an argument to output in a tidy (\texttt{longer=TRUE}) format, and the example below illustrates how to spread contingency table statistics by class:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth, }
                 \AttributeTok{longer =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Other"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{spread}\NormalTok{(Class, Value) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{relocate}\NormalTok{(Average, }\AttributeTok{.after=}\NormalTok{dplyr}\SpecialCharTok{::}\FunctionTok{last\_col}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate\_if}\NormalTok{(rlang}\SpecialCharTok{::}\NormalTok{is\_double, signif, }\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 14 x 7
##    Statistic                 `0`    `1`    `2`    `3` unrelated Average
##    <chr>                   <dbl>  <dbl>  <dbl>  <dbl>     <dbl>   <dbl>
##  1 Balanced Accuracy       0.86   0.88   0.9   0.84       0.91    0.88 
##  2 Detection Prevalence    0.19   0.24   0.17  0.083      0.31    0.2  
##  3 Detection Rate          0.17   0.2    0.14  0.021      0.28    0.16 
##  4 F1/Dice                 0.8    0.82   0.83  0.38       0.88    0.74 
##  5 FDR                     0.14   0.17   0.17  0.75       0.089   0.26 
##  6 FNR                     0.25   0.19   0.17  0.25       0.15    0.2  
##  7 FOR                     0.069  0.064  0.033 0.0076     0.071   0.049
##  8 FPR/Fallout             0.036  0.056  0.033 0.064      0.042   0.046
##  9 N                      64     72     48     8         96      58    
## 10 NPV                     0.93   0.94   0.97  0.99       0.93    0.95 
## 11 PPV/Precision           0.86   0.83   0.83  0.25       0.91    0.74 
## 12 Prevalence              0.22   0.25   0.17  0.028      0.33    0.2  
## 13 Sensitivity/Recall/TPR  0.75   0.81   0.83  0.75       0.85    0.8  
## 14 Specificity/TNR         0.96   0.94   0.97  0.94       0.96    0.95
\end{verbatim}

@@

Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020} detect shared IBD segments between individuals. The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and converts that shared cM to a kinship coefficient. In addition to inferred segments, these functions can estimate ``truth'' kinship from simulated IBD segments \citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise IBD segments from IBD inference tools and from simulated IBD segments. The \texttt{read\_map()} function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the \texttt{ibd2kin()} function. See \texttt{?read\_ibd} and \texttt{?read\_map} for additional details on expected format.

The \texttt{read\_ibd()} function reads in the pairwise IBD segment format. Input to this function can either be inferred IBD segments from hap-IBD (\texttt{source="hapibd"}) or simulated segments (\texttt{source="pedsim"}). The first example below uses data in the \texttt{hap-ibd} output format:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hapibd\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
                               \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
\NormalTok{hapibd\_seg }\OtherTok{\textless{}{-}} \FunctionTok{read\_ibd}\NormalTok{(hapibd\_filepath, }\AttributeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
\NormalTok{hapibd\_seg}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 3,954 x 6
##    id1               id2                 chr     start       end length
##    <chr>             <chr>             <dbl>     <dbl>     <dbl>  <dbl>
##  1 testped1_g1-b1-s1 testped1_g3-b1-i1     1 197661576 234863602   47.1
##  2 testped1_g2-b2-i1 testped1_g3-b1-i1     1 197661576 231017545   39.8
##  3 testped1_g3-b1-i1 testped1_g3-b2-i1     1 197661576 212799139   20.3
##  4 testped3_g1-b1-s1 testped3_g3-b2-i1     1   2352146  10862397   17.7
##  5 testped3_g2-b2-i1 testped3_g3-b2-i1     1   2352146  10862397   17.7
##  6 testped1_g1-b1-s1 testped1_g2-b1-i1     1   3328659  64123868   86.4
##  7 testped1_g1-b1-s1 testped1_g3-b1-i1     1   3328659  33476811   51.2
##  8 testped1_g2-b2-s1 testped1_g3-b2-i1     1   5003504  32315147   45.9
##  9 testped2_g1-b1-i1 testped2_g3-b1-i1     1 240810528 248578622   15.9
## 10 testped2_g1-b1-i1 testped2_g2-b2-i1     1 241186056 249170711   15.5
## # ... with 3,944 more rows
\end{verbatim}

In order to translate the shared genomic cM length to a kinship coefficient, a genetic map must first be read in with \texttt{read\_map()}. Software for IBD segment inference and simulation requires a genetic map. The map loaded for kinship estimation should be the same one used for creating the shared IBD segment output. The example below uses a minimal genetic map that ships with \texttt{skater}:
